\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{SAVNav: Socially-Aware Audio Visual Navigation\\
% {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
% should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
% \and
% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% \IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{METHODOLOGY}\label{methodology}

\subsection{SYSTEM OVERVIEW \& PROBLEM
FORMULATION}\label{system-overview--problem-formulation}

We consider a mobile robot operating in a dynamic, semi-structured
indoor environment populated by human agents. The robot is equipped with
an RGB-D camera and a 4-channel microphone array.

Coordinate Systems: We define the World Frame \(\mathcal{W}\) with the
origin at the scene center and the Ego Frame \(\mathcal{E}\) attached to
the robot\textquotesingle s geometric center. Let
\(\mathbf{x}_t = [x, y, z, \psi]^T\) denote the robot\textquotesingle s
pose in \(\mathcal{W}\) at time \(t\).

Problem Definition: Given a natural language instruction \(\mathcal{I}\)
(e.g., \emph{"Find the person talking"}), the objective is to generate a
sequence of control actions \(a_{t} \in \mathcal{A}\) to navigate
towards the target location \(\mathbf{g}^*\) while minimizing the
traversal time and ensuring social compliance. This requires handling
two types of entities: Observable Entities (within the
camera\textquotesingle s Field-of-View, FoV) and Occluded Acoustic
Sources (Non-Line-of-Sight, NLOS).

\subsection{MULTI-MODAL PERCEPTION WITH
UNCERTAINTY}\label{multi-modal-perception-with-uncertainty}

Our perception stack processes visual and auditory streams
asynchronously to extract semantic and spatial information, explicitly
modeling the uncertainty inherent in each modality.

\subsubsection{Visual Perception}\label{visual-perception}

The visual module runs at camera frame rate (\(\sim\) 20Hz). We employ
YOLOv11 {[}@YOLOERealtimeSeeing2025{]} for 2D object detection and
segmentation. For each detected bounding box, we compute the 3D centroid
\(\mathbf{p}_{vis}^{\mathcal{E}}\) in the Ego Frame using depth data.\\
To account for sensor noise and partial occlusions, we model the
positional uncertainty \(\sigma_{p}\) as:

\[\sigma_{p} = k_d \cdot z^2 + k_{occ} \cdot \left(1 - \frac{N_{mask}}{N_{box}}\right)\]

where \(z\) is the depth, \(k_d\) captures the depth
sensor\textquotesingle s quadratic error characteristics, and the second
term penalizes occlusion based on the mask-to-box ratio with coefficient
\(k_{occ}\). We extract semantic features \(\mathbf{f}_{vis}\) using
ImageBind {[}@ImageBindOneEmbedding2023{]} and appearance features
\(\mathbf{f}_{reid}\) for temporal tracking.

\subsubsection{Audio Perception}\label{audio-perception}

The audio module processes 4-channel ambisonic signals in a sliding
window (5s duration, 0.5s stride). We utilize Embed-ACCDOA
{[}@OpenvocabularySoundEvent2025{]} for Sound Event Localization and
Detection (SELD).\\
For each frame, the model predicts the Direction-of-Arrival (DoA) vector
\(\mathbf{v}_{doa} \in \mathbb{R}^3\) and a semantic label. The angular
uncertainty \(\sigma_{ang}\) is dynamically scaled by the detection
confidence \(c\):

\[\sigma_{ang} = \sigma_{base} \cdot (1 + k_{unc} \cdot (1 - c))\]

where \(\sigma_{base}\) is the empirical error margin of the SELD model.
Unlike vision, audio lacks direct distance information; thus, we model
the source position as a probability cone
\(\mathcal{C}_{audio}(\mathbf{x}_t, \mathbf{v}_{doa}, \sigma_{ang})\)
originating from the robot.

\subsection{SAVMAP: SPATIO-TEMPORAL
MAPPING}\label{savmap-spatio-temporal-mapping}

We introduce SAVMap, a layered grid map (0.05m resolution) that
integrates static semantics with dynamic audio-visual entities to
support socially-aware navigation.

\subsubsection{Entity Tracking and
Association}\label{entity-tracking-and-association}

We maintain a list of tracked entities. For visual observations, we
employ a Hungarian matching algorithm based on a hybrid cost matrix
\(C_{ij}\):

\[C_{ij} = w_{pos} \cdot \|\mathbf{p}_i - \mathbf{p}_j\| + w_{emb} \cdot (1 - \cos(\mathbf{f}_i, \mathbf{f}_j))\]

For audio observations, we perform cross-modal association. If an audio
source\textquotesingle s DoA aligns with a visual entity within the
angular threshold \(\sigma_{ang}\) and shares semantic consistency
(e.g., "speech" vs. "person"), the audio event is anchored to the visual
target.

\subsubsection{Topology-Aware Hallucinated
Momentum}\label{topology-aware-hallucinated-momentum}

A key contribution of our work is handling unanchored audio sources
(i.e., heard but not seen). To anticipate NLOS risks (e.g., a person
approaching from around a corner), we propose a Hallucinated Momentum
Injection mechanism:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Ray-Casting: We project a ray along the unanchored DoA until it
  intersects with the static map boundaries.
\item
  Topological Adsorption: The intersection point is snapped to the
  nearest topological node (e.g., a doorway or corner), denoted as the
  hallucinated position \(\mathbf{p}_{hall}\).
\item
  Virtual Velocity: We assign a virtual velocity vector
  \(\mathbf{v}_{virt}\) to this entity, pointing directly towards the
  robot, simulating a worst-case collision scenario.
\end{enumerate}

\subsubsection{Social Cost Layer}\label{social-cost-layer}

We construct a unified cost map \(M_{cost}\) to guide the planner.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Visible Humans: For tracked visual entities, we generate a Gaussian
  cost field. If the human is moving, the Gaussian is anisotropic,
  elongated along their velocity vector.
\item
  NLOS Audio Sources: For hallucinated entities, we generate a
  Lance-shaped Risk Field aligned with \(\mathbf{v}_{virt}\) to penalize
  the path directly in front of the hidden source:
\end{enumerate}

\[J_{audio}(\mathbf{x}) = A \cdot \exp \left( - \left( \frac{d_{long}^2}{2\sigma_{long}^2} + \frac{d_{lat}^2}{2\sigma_{lat}^2} \right) \right)\]

where \(d_{long}\) and \(d_{lat}\) are longitudinal and lateral
distances to the source in the frame aligned with \(\mathbf{v}_{virt}\).
By setting \(\sigma_{long} \gg \sigma_{lat}\), this cost forces the
planner to take wide turns at corners ("cutting the corner wide"),
effectively implementing a "Hearing before Seeing" behavior.

\subsection{SAVNAV: SOCIALLY-AWARE
PLANNING}\label{savnav-socially-aware-planning}

To handle the ambiguity of audio data, we design a dual-mode planner
that switches between Precision Navigation and Active Exploration.

\subsubsection{State Arbitration}\label{state-arbitration}

The system state is determined by the target entity\textquotesingle s
status:

\begin{itemize}
\item
  State A (Precision Nav): Triggered when the target is visually
  anchored or statically defined (e.g., "Door").
\item
  State B (Active Exploration): Triggered when the target is only
  detected via audio (Unanchored).
\end{itemize}

\subsubsection{State A: Risk-Aware Path
Generation}\label{state-a-risk-aware-path-generation}

In this state, the goal \(\mathbf{g}^*\) is deterministic. We employ the
Fast Marching Method (FMM) to solve the Eikonal equation over the speed
field \(S(\mathbf{x}) = S_{max} \cdot (1 - M_{cost}(\mathbf{x}))\). The
gradient descent on the resulting arrival time map \(T(\mathbf{x})\)
yields a smooth, global path that naturally skirts high-social-cost
regions.

\subsubsection{State B: Information-Theoretic Active
Exploration}\label{state-b-information-theoretic-active-exploration}

When the target position is uncertain (audio-only), the robot must
actively reduce entropy. We define the Region of Interest (ROI)
\(\Omega_{ROI}\) as the intersection of the audio uncertainty cone
\(\mathcal{C}_{audio}\) and the navigable space.\\
We select the Next-Best-View (NBV) pose \(\mathbf{p}_{nbv}\) by
maximizing an Information-Theoretic Utility function \(U(\mathbf{p})\):

\[U(\mathbf{p}) = \underbrace{\text{Area}(\text{FOV}(\mathbf{p}) \cap \Omega_{ROI} \setminus \mathcal{V}_{hist})}_{\mathcal{I}_{gain}} \cdot \underbrace{\exp(-\lambda \cdot \|\mathbf{x}_t - \mathbf{p} \|)}_{\mathcal{C}_{cost}}\]

where \(\mathcal{I}_{gain}\) represents the expected volumetric
information gain (unobserved ROI area visible from \(\mathbf{p}\)), and
\(\mathcal{C}_{cost}\) penalizes travel distance. \(\mathcal{V}_{hist}\)
tracks previously observed regions to prevent redundant exploration.\\
Once the robot reaches \(\mathbf{p}_{nbv}\) or the target enters the FoV
during transit, the system transitions to State A for final approach.

\subsubsection{Reactive Control}\label{reactive-control}

The planned path is executed by a local controller. To ensure safety, an
Emergency Brake is triggered if the Time-To-Collision (TTC) with any
dynamic obstacle (visual or hallucinated) drops below 1.0s.

\end{document}
