\section{Related Work}
\label{sec:related}

\subsection{Vision-Based Social Navigation}
Traditional and recent learning-based SocialNav systems strive to navigate safely and politely around humans \cite{PrinciplesGuidelinesEvaluating2025, Habitat30Cohabitat2023}. A common theme across these systems is adjusting the path to not only avoid physical collisions but also respect implicit social norms. These approaches primarily rely on LOS sensors, such as RGB-D cameras or LiDAR, to detect pedestrians and forecast their future trajectories \cite{SocialNavTrainingHumanInspired2026}. For instance, Habitat 3.0 \cite{Habitat30Cohabitat2023} provides a co-habitat simulation for humanoid-robot collaboration, driving socially compliant visual policies. Falcon \cite{CognitionPrecognitionFutureaware2025} introduces a proactive reinforcement learning framework that predicts future human movements to penalize actions that might block human paths. 

Despite their success in visible spaces, these vision-centric policies suffer from a fundamental constraint: they completely ignore NLOS risks. Without the ability to perceive beyond the visual field, a purely visual agent cannot anticipate a dynamic pedestrian approaching from a blind corner or a group of people engaged in conversation behind a wall. This often leads to abrupt stops, potential collisions, and severe violations of social boundaries when the unexpected human suddenly enters the LOS.

\subsection{Audio-Visual Navigation}
To overcome the limitations of restricted visual fields, AVNav equips embodied agents with auditory perception, offering omnidirectional and space-penetrating cues. Early AVNav research \cite{LookListenAct2020, SoundSpaces20Simulation2022} demonstrated that sound can effectively guide an agent to an acoustic target even in unexplored environments, substantially broadening the scope of pure visual object navigation \cite{ObjectNavRevisitedEvaluation2020}. 

As the task has deepened, researchers have explored more complex and realistic scenarios. The semantic audio-visual navigation (SAVi) task \cite{SemanticAudiovisualNavigation2021} removes the assumption of continuous sound emission, requiring agents to discover targets from transient semantic acoustic events. Recent works expand this to highly dynamic or noisy environments, tracking moving targets \cite{CatchMeIf2023} and tackling multi-source interference \cite{AudiovisualNavigationNoisy2025, PaySelfattentionAudiovisual2022}. Furthermore, several methods incorporate cross-modal attention \cite{OmnidirectionalInformationGathering2023} and leverage foundation models \cite{RILAReflectiveImaginative2024, CAVENEmbodiedConversational2024} to enhance semantic reasoning and zero-shot exploration.

While these AVNav approaches excel at localizing an acoustic target, they share a critical blind spot concerning social context. They typically treat non-target acoustic events (e.g., human footsteps or conversations) either as irrelevant background noise to be filtered out or simply as alternative navigation goals \cite{AudiovisualNavigationNoisy2025}. Consequently, they fail to model these social acoustic events as spatial boundaries, leaving the agent prone to disturbing stationary conversations or colliding with NLOS dynamic pedestrians while ruthlessly pursuing its target.

\subsection{Cross-Modal Spatial Memory and Mapping}
As navigation systems scale to long-horizon and complex tasks, there has been a significant push toward structured scene representations. Purely visual architectures such as 3D Dynamic Scene Graphs \cite{3DDynamicScene2020} or open-vocabulary mapping frameworks \cite{ConceptFusionOpensetMultimodal2023, BareQueriesOpenvocabulary2025} map geometric features to high-level semantic nodes for complex spatial reasoning. For AVNav, memory-augmented learning \cite{EchoenhancedEmbodiedVisual2023} and acoustic waypoint policies (AV-WaN \cite{LearningSetWaypoints2021}) have been used to tackle forgetting in long paths and bridge low-level control with high-level topological planning.

Although these spatial mappings provide actionable topology \cite{LearningSetWaypoints2021}, they remain structurally unequipped for the SAVNav task. Purely visual scene graphs can only register entities once they enter the agent's field of view \cite{3DDynamicScene2020}, missing hidden objects entirely. Similarly, existing AVNav spatial memories merely accumulate historical trajectory features to find the target \cite{EchoenhancedEmbodiedVisual2023, LearningSetWaypoints2021}, rather than treating the audio as a map-wide constraint. 

Our proposed acoustic-to-spatial social mapping module, SAVMap, fills this crucial gap. Instead of relying purely on direct sight or treating audio events solely as goals, SAVMap records transient acoustic events and anchors them as inferred entities within the physical topological layout. It explicitly projects these non-visual inferences into a continuous social cost field, successfully transforming fleeting multidirectional sounds into persistent, geometry-aware social boundaries for proactive, socially compliant planning.
